---
title: "Political Party Identification & Environmentalism"
author: "Kirsten White"
output: 
  html_document: 
    df_print: kable
    theme: yeti
---

```{r setup, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Attach Packages

library(tidyverse)
library(here)
library(broom)
library(janitor)
```

```{r, include = FALSE}
# Read in the data 

conservation_survey <- read_csv(here("data", "conservation_survey.csv"))

```

```{r, include = FALSE}
# Create a subset containing only ENV_ID and POL_OR columns
conservation_subset <- conservation_survey %>%
  select(ENV_ID, POL_OR) %>%
  mutate(POL_PARTY = case_when
         (POL_OR %in% c("1","2", "3") ~ "Conservative", 
           POL_OR %in% c("5", "6", "7") ~ "Liberal")) %>%
          drop_na() %>% # conservative 1-3; liberal 5-7; exclude 4, 8, 9
  mutate(ENV_ID = case_when(
         ENV_ID == "1" ~ "Yes", 
         ENV_ID == "2" ~ "No", 
         ENV_ID == "3" ~ "Unsure"))
```

```{r, include = FALSE}
# Create a finalized table showing the counts and proportions of liberal and conservative respondents who responded yes, no, and unsure

# Find the counts 
conservation_subset %>%
  count(POL_PARTY) %>%
  arrange(-n)

# Find counts by political party affiliation and response (to environmentalist q) 
conservation_subset %>%
  count(POL_PARTY, ENV_ID)

# Return counts (from above) in a contingency table format 
environmental_counts <- conservation_subset %>%
  tabyl(POL_PARTY, ENV_ID)

# Show percentages 
environmental_proportions <- environmental_counts %>%
  adorn_percentages() %>%
  adorn_pct_formatting((digits = 2)) %>%
  adorn_ns()

# Convert to contingency table 

environmental_table <- environmental_counts %>%
  column_to_rownames(var = "POL_PARTY")

#prop_table <- environmental_proportions %>%
 # column_to_rownames(var = "POL_PARTY")

# Chi-squared - asking: is identifying as an environmentalist truly independent of the political party one identifies with 
environ_x2 <- chisq.test(environmental_table)
environ_x2

# Write out chi-sq info in tidy format (in order to refer to the values in-line)
environ_tidy <- tidy(environ_x2)

```

---

## Introduction

In this report, we will explore the following question: is the
an association between political affiliation (conservative or liberal) and whether or not a person considers themselves an environmentalist? To explore this question, responses from a survey conducted by Chelsea Batavia and Andrews Forest to explore the "influences on charitable giving for conservation" was used as the data. The survey, conducted in 2017, included responses from 1,331 participants. 

To explore this relationship, a contingency table was created with the counts and proportions of participant responses to the question above. Then, a chi-square test for independence was performed to assess the association between political party and environmentalism. 

---

## Findings

```{r}
# Print out a finalized version of this contingency table 

environmental_proportions
```
**Table 1**: The majority of conservative identifying participants responded "No" they do not consider themselves environmentalists (50.24%). Conversely, the majority of liberal identifying participants responded "Yes" they do consider themselves environmentalists (53.73%). 


---

## Conclusion

Following the analysis, a significant association was found between the political party a person affiliates with and whether or not they consider themselves an environmentalist: ($\chi$^2^(`r environ_tidy$parameter`) = `r round(environ_tidy$statistic, 2)`, *p* `r ifelse(environ_tidy$p.value < 0.001, paste("<0.001"), paste("=", environ_tidy$p.value))`). Conservative-identifying people more often did not identify as environmentalists while liberal-identifying people more often did identify as environmentalists. 

**Data Citation**: Batavia, C. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. <https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c> (Accessed 2021-11-23).
